### 2021-07-13 submission.tex
submission

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{alltt}
\usepackage[top=2.5cm, left=3cm, right=3cm, bottom=4.0cm]{geometry}

@@

	{\Large Karol Borkowski \\ \emph{nethz} Username: kborkowski \\ Student ID: 00000000}\\[0.5cm] 
	\vfill
	{\large \textbf{Collaborators:} \\
	Other student 1 \\
	Other student 2}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[scale=1.5]{figures/backprop.jpg}
    \caption{The computation graph of f.}
    \label{fig:roller-coaster}
\end{figure}

@@

\section*{Question 4} 
\begin{enumerate}[label = (\alph*)]
    \item
    \begin{enumerate}[label = (\roman*)]
    \item
    Let's define an auxiliary function $J_w = \begin{cases}
      1, & \text{if} \; w \; is \; drawn \\
      0, & \text{otherwise}
    \end{cases}$

@@

    Finally: $X = |V| \cdot (\frac{|V|-1}{|V|})^n$

@@

    $|\bigcup_{i=1}^{|V|}P(!w_i)| = \sum_{J \in \{1,...,|V|\}} (-1)^{|J|+1}|\bigcup_{j \in J}P(!w_i)|) = $

@@

    \item
    TO DO

@@

\section*{Question 5} 
\begin{enumerate}[label = (\alph*)]
    \item
    \begin{enumerate}[label = (\roman*)]
    \item
    G = <V, E, W>

@@

    W - ???

@@

@@

\clearpage
\newpage

\begin{table}[h]
        \centering
        \fontsize{10}{10}\selectfont
        \renewcommand{\arraystretch}{1.2} % vertical padding
        \setlength{\tabcolsep}{0.5em} % for the horizontal padding
        \begin{tabular}{l|l|l|l}
        \textbf{number} & \textbf{sample strings} & \textbf{accepted} & \textbf{weight} \\ \hline
        1 & educational is this not &  &  \\
        2 & is this assignment educational &  &  \\
        3 & not educational is not educational &  &  \\
        4 & this assignment is not educational &  &  \\
        5 & is this assignment educational &  &  \\
        6 & this assignment course is educational &  &  \\
        7 & is this assignment not educational &  &  \\
        8 & this assignment not &  &  \\
        9 & this course assignment is not educational &  &  \\
        10 & this course is not not educational &  &  \\
        11 & not educational is this &  &  \\
        12 & course assignment is not educational &  &  \\
        13 & not this assignment is educational &  &  \\
        14 & not not not educational &  &  \\
        14 & is this course assignment not educational &  &  \\
        15 & course assignment is this &  &  \\
        16 & this course is interesting &  &  \\
        17 & this course assignment not educational &  & 
        \end{tabular}
        \caption{Some strings from $\mathcal{Y}_{\geq 2, \leq 6}$}
        \label{tab:wfst_strings}
        \end{table}

\begin{figure}[!h]
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/wfst_charts_1.pdf}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 0 to 3; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_1}
    \end{figure}

    \begin{figure}[!h]
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/wfst_charts_2.pdf}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 4 to 7; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_2}
    \end{figure} 
\end{document}

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{alltt}
\usepackage{float}
\usepackage[top=2.5cm, left=3cm, right=3cm, bottom=4.0cm]{geometry}

@@

	{\Large Karol Borkowski \\ \emph{nethz} Username: kborkowski \\ Student ID: 20-905-501}\\[0.5cm] 
	\vfill
	{\large \textbf{Collaborators:} \\
	Sven Kohler}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[scale=1.5]{figures/backprop.jpg}
    \caption{The computation graph of f.}
\end{figure}

@@

\clearpage
\newpage
\section*{Question 4} 
\begin{enumerate}[label = (\alph*)]
    \item
    \begin{enumerate}[label = (\roman*)]
    \item
    Let's define an auxiliary function $J_w = \begin{cases}
      1, & \text{if} \; w \; is \; drawn \\
      0, & \text{otherwise}
    \end{cases}$

@@

    Finally: $X = |V| \cdot (1 - \frac{|V|-1}{|V|})^n$

@@

    $|\bigcup_{i=1}^{|V|}P(!w_i)| = \sum_{J \in \{1,...,|V|\}} (-1)^{|J|+1}|\bigcup_{j \in J}P(!w_i)|) $

@@

    \item
    Probability that the word "work" is selected in n draws is equal to 1 - probability that it is not selected.

    $P('work' 
   \; selectes) = 1 - P('work' \; not \; selected) = 1 - (\frac{|V|-1}{|V|})^n$

   To find the expected number of draws let's put 0.95 for P and solve the above equation for variable n:

   $1 - (\frac{|V|-1}{|V|})^n \geq 0.95$

   $(\frac{|V|-1}{|V|})^n \leq 0.05$

   $n \leq log_{\frac{|V|-1}{|V|}}(0.05)$

@@

\clearpage
\newpage
\section*{Question 5} 
\begin{enumerate}[label = (\alph*)]
    \item
    \begin{enumerate}[label = (\roman*)]
    \item
    G = $<V, E, W>$

@@

    W = $score_{w} - score_{max}, \; for \; each \; w \; \in \; vocabulary $

    To solve the decoding problem we need to find the highest scoring path, what is equivalent to finding the longest path on a graph with negative weights. The negativity of the weights is ensured using the above transformation for W. 
    The following algorithm is guaranteed to work only in the case of a directed trelis.

@@

    Used semiring:

    In case of positive weights and the shortest path: $<R^+ \bigcup \{\inf\}, min, +, \inf, 0>$

    The case of the longest path and negative weights is shown in thpoint ii).

    \item
    $<R^- \bigcup \{- \inf\}, max, +, -\inf, 0>$

    \item
    $<R^+ \bigcup \{\inf\}, max, min, 0, \inf>$

@@

\clearpage
\newpage
\part{Course Assignment Episode 2}

\section*{Question 1} 
\begin{enumerate}[label = (\alph*)]
    \item
    $<N, \Sigma, R, S>$

    N = {NP, VP, PP, V, NP, Det, N, P}

    $\Sigma$ = {I, draw, a, man, with, pencil, hit, ball, an, umbrella, glasses}

    R:

    $S \to NP \;\; VP$

    $NP \to Det \;\; N$

    $NP \to Np \;\; PP$

    $V \to VP \;\; PP$

    $VP \to V \;\; NP$

    $PP \to P \;\; NP$

    $NP \to J \; | \; glasses$

    $V \to draw \; | \; hit$

    $Det \to a \; | \; an$

    $N \to man \; | \; pencil \; | \; ball \; | \; umbrella$

    $P \to with$

    \item
    $S \to NP \;\; VP$: 1

    $NP \to Det \;\; N$: 1/2

    $NP \to Np \;\; PP$: 1/7

    $NP \to J$: 2/7

    $NP \to glasses$: 1/14

    $V \to draw$: 1/2

    $V \to hit$: 1/2

    $Det \to a$: 6/7

    $Det \to an$: 1/7

    $N \to man$: 4/7

    $N \to pencil$: 1/7

    $N \to ball$: 1/7

    $N \to umbrella$: 1/7

    $P \to with$: 1

    $PP \to P \;\; NP$: 1

    $VP \to VP \;\; PP$: 1/3

    $VP \to V \;\; NP$: 2/3

    \item
    To capture different expansions of NP, we introduce the the following rules:

    $NP \to N$ and $N \to J$

    The new PCFG is:

    $S \to NP \;\; VP$: 1

    $NP \to Det \;\; N$: 1/2

    $NP \to Np \;\; PP$: 1/7

    $NP \to N$: 5/14

    $V \to draw$: 1/2

    $V \to hit$: 1/2

    $Det \to a$: 6/7

    $Det \to an$: 1/7

    $N \to man$: 4/12

    $N \to pencil$: 1/12

    $N \to ball$: 1/12

    $N \to umbrella$: 1/12

    $N \to J$: 5/12

    $P \to with$: 1

    $PP \to P \;\; NP$: 1

    $VP \to VP \;\; PP$: 1/3

    $VP \to V \;\; NP$: 2/3

	\begin{figure}[H]
	\item
    \centering
    \includegraphics[scale=1.5]{figures/tree.jpg}
    \caption{The parse tree for the sentence “I hit a man with glasses” with 		lexicalized rules.}
	\end{figure}

	\item
	The time complexity of CKY algorithm is $O(n^3 \cdot |G|)$, 
	where n is the length of the parsed string and $|G|$ is the size of the CNF grammar G.

	The modifications from parts c) and d) increases the number of the grammar rules $|G|$ and therefore the time of the algorithm execution. Nevertheless, the order expressed by the big-O notation is the same.

	The space complexity of the CKY algorithm is determined by the size of the used matrix, which is: $O(n^2)$

\clearpage
\newpage
\section*{Question 2} 
\begin{enumerate}[label = (\alph*)]
    \item
    Let's create a lexicalized production for each $\psi(i \to j)$,

    where i $<$ j:

    $X(w_i) \to X(w_i)X(w_j) \; : \; \psi(i \to j)$,

    where i $>$ j:

    $X(w_i) \to X(w_j)X(w_i) \; : \; \psi(i \to j)$

    and for each $\psi(root \to j)$:

    $S \to X(w_j) \; : \; \psi(root \to j)$

    \item
    Let's consider all possible dependency parses and their associated
derivations in the lexicalized CFG:

	$\psi_\tau(y^1) = \psi_\tau(S \to X(They)) + \psi_\tau(X(They) \to X(They X(fish))) = \psi(root \to 1) + \psi(1 \to 2) = \Psi(1 \to 2)$

		$\psi_\tau(y^2) = \psi_\tau(S \to X(fish)) + \psi_\tau(X(fish) \to X(They X(fish))) = \psi(root \to 2) + \psi(2 \to 1) = \Psi(2 \to 1)$

\clearpage
\newpage
\section*{Question 4}

\begin{enumerate}[label = (\alph*)]

\begin{table}[h]
\item
        \centering
        \fontsize{10}{10}\selectfont
        \renewcommand{\arraystretch}{1.2} % vertical padding
        \setlength{\tabcolsep}{0.5em} % for the horizontal padding
        \begin{tabular}{l|l|l|l}
        \textbf{number} & \textbf{sample strings} & \textbf{accepted} & \textbf{weight} \\ \hline
        1 & educational is this not & NO &  \\
        2 & is this assignment educational & NO  &  \\
        3 & not educational is not educational & YES  & 12  \\
        4 & this assignment is not educational &  YES & 9  \\
        5 & is this assignment educational & NO &  \\
        6 & this assignment course is educational & NO  &  \\
        7 & is this assignment not educational & YES & 15  \\
        8 & this assignment not & YES & 8  \\
        9 & this course assignment is not educational & YES & 12  \\
        10 & this course is not not educational & YES & 21 \\
        11 & not educational is this & YES & 11 \\
        12 & course assignment is not educational & YES & 10 \\
        13 & not this assignment is educational & NO &  \\
        14 & not not not educational & YES & 25 \\
        14 & is this course assignment not educational & YES & 25 \\
        15 & course assignment is this & 15 & 8 \\
        16 & this course is interesting & NO &  \\
        17 & this course assignment not educational & YES & 14
        \end{tabular}
        \caption{Some strings from $\mathcal{Y}_{\geq 2, \leq 6}$}
        \label{tab:wfst_strings}
        \end{table}

\begin{figure}[H]
		\item
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/floyd1.jpg}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 0 to 3; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_1}
    \end{figure}

    \begin{figure}[H]
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/floyd2.jpg}
        % \includegraphics[width=.45\linewidth]{./fig/C1000.png}
        \caption{Floyd-Warshall algorithm, iteration 4 to 7; left column matrix should contain weights after iteration n; right column matrix should be iteratively filled for backtracking each path}
        \label{fig:wfst_charts_2}
    \end{figure}

	\item
	The number of iterations is bound by the number of nodes. It terminates after n=$|V|$

	\item    
	Time complexity: three for loops $=> \; O(n^3)$

	Space complexity: size of the matrices = $ 2 \cdot n^2 \; => \; O(n^2)$

	Backtracking: Maximal when all edges are included into the path $=> \; O(|E|)$

    \end{enumerate}
\end{document}
